---
title: 'DRAFT: Lab 3: Reducing Crime'
author: "C. Akkineni, A. Thorp, K. Hanna"
date: "November 27, 2018"
output:
    pdf_document:
    toc: true
    toc_depth: 2
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\tableofcontents 
<!--
\listoffigures
\listoftables
-->
\newpage

# Stage 1: Draft Report
```{r}
crime = read.csv('crime_v2.csv')
# Delete the 6 empty rows at the end
crime[92:100,]
crime = crime[1:91, ]

# Convert columns to factors and logical.
crime$county = as.factor(crime$county)
crime$year = as.factor(crime$year)
crime$west = as.logical(crime$west)
crime$central = as.logical(crime$central)
crime$urban = as.logical(crime$urban)

# Fix prbconv, convert from factor to numeric
summary(crime$prbconv)
crime$prbconv = as.numeric(crime$prbconv)

# county 193 is duplidated, remove one
crime[crime$county == 193, ]
crime = crime[-c(89), ]

summary(crime)

```

# Preliminary Infomations
## Variables:

1. Target
+ crmrte  

2. Label
+ county

3. Segregates: (my own word, just things that can segregate the data).  Counties can belong to 0 or more from west, central and urban.
+ density (likely related to others, especially urban)  
+ west  
+ central  
+ urban   
  

4. Cost of doing crime:  
+ prbconv  
+ prbpris  
+ avgsen  
+ prbarr  
+ polpc (likely related to prbconv)  



```{r} 
plot(crime$crmrte, ylab = 'crimes committed per person')
```

```{r}
hist(crime$crmrte, xlab = 'crimes committed per person', main = 'Histogram of crimes committed per person')
```

```{r}
model1 = lm(crmrte ~ prbarr + polpc + density, data = crime)
(model1$coefficients)

```


## Steps for evaluating variables
Leverage (and Influence if required)  
Goodness-of-Fit : AIC  
Endoginaity  
Omitted variable bias  
MSE  
E[theta hat] = theta

```{r}
sum(c(crime$urban, crime$west))
```


# 1. What do you want to measure? Make sure you identify variables that will be relevant to the concerns of the political campaign.

# 2. What transformations should you apply to each variable? This is very important because transformations can reveal linearities in the data, make our results relevant, or help us meet model assumptions.

# 3. Are your choices supported by EDA? You will likely start with some general EDA to detect anomalies (missing values, top-coded variables, etc.). From then on, your EDA should be interspersed with your model building. Use visual tools to guide your decisions.

# 4. What covariates help you identify a causal effect? What covariates are problematic, either due to multicollinearity, or because they will absorb some of a causal effect you want to measure?

