---
title: "Kevin's Sandbox"
author: "Kevin Hanna"
date: "November 23, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(knitr)
library(kableExtra)
library(car)

codebook <- read.csv('codebook.csv')
crime <- read.csv('crime_v2.csv')

# Convert columns to factors and logical.
crime$county <- as.factor(crime$county)
crime$year <- as.factor(crime$year)
crime$west <- as.logical(crime$west)
crime$central <- as.logical(crime$central)
crime$urban <- as.logical(crime$urban)

# Create a log of the dependent variable
crime$logcrmrte <- log(crime$crmrte)

# Delete the 6 empty observations at the end, including the row with the apostrophe.
# We can use complete.cases to do this as these 6 observations are the only incomplete observations. 
crime = crime[complete.cases(crime), ]

# Fix prbconv which is a factor rather than numeric due to the apostrophe
# Convert from factor to numeric 
crime$prbconv = as.numeric(as.character(crime$prbconv))

# county 193 is duplidated, remove one
crime = crime[!duplicated(crime), ]

```

# Preliminary Infomations (not intended to be left in)

## From the assignment:
+ 1. What do you want to measure? Make sure you identify variables that will be relevant to the concerns of the political campaign.
+ 2. What transformations should you apply to each variable? This is very important because transformations can reveal linearities in the data, make our results relevant, or help us meet model assumptions.
+ 3. Are your choices supported by EDA? You will likely start with some general EDA to detect anomalies (missing values, top-coded variables, etc.). From then on, your EDA should be interspersed with your model building. Use visual tools to guide your decisions.
+ 4. What covariates help you identify a causal effect? What covariates are problematic, either due to multicollinearity, or because they will absorb some of a causal effect you want to measure?

## Variables:


### 1. Target
+ crmrte  

### 2. Label
+ county

### 3. Geographic:
+ density (likely related to others, especially urban)  
+ west  
+ central  
+ urban   

Correlation between logcrmrte and urban: 0.491 and with density 0.633.  
Correlation between urban and density is 0.820  
Correlation between logcrmrte and west is -0.414  west is also negatively correlated with density.  

I think density is an important variable (more so than urban).  This would be logical as low income housing is often high-density.  


```{r}
# Geographic
#foo2 = lm(crmrte ~ urban + central + west + density, data = crime)
#foo2$coefficients
#vcov(foo2)

foo2log = lm(logcrmrte ~ urban + central + west + density, data = crime)
foo2log$coefficients
#vcov(foo2log)

foo2rows = c("logcrmrte", "crmrte", "urban", "central", "west", "density")

round(cor(crime[foo2rows]), 3)
scatterplotMatrix(crime[,foo2rows], diagonal = "histogram")

```


### 4. Cost of doing crime:  
#### Probabilities:
+ prbconv  
+ prbpris  
+ prbarr  

Both prbarr and prbconv are negativly correlated to logcrmrte (-0.473 and -0.447 respectively).  prbconv is less reliable (unless we can explain the > 1 values.)  


```{r}
# Probabilities

#foo1 = lm(crmrte ~ prbarr + prbconv + prbpris, data = crime)
#foo1$coefficients
#vcov(foo1)

foo1log = lm(logcrmrte ~ prbarr + prbconv + prbpris, data = crime)
foo1log$coefficients
#vcov(foo1log)

foo1rows = c("logcrmrte", "crmrte", "prbarr", "prbconv", "prbpris")

round(cor(crime[foo1rows]), 3)
scatterplotMatrix(crime[,foo1rows], diagonal = "histogram")

```


#### Sentence and police
+ avgsen  
+ polpc (likely related to prbconv)  

polpc has a huge correlation, it makes sense, but it's still so high we should be very cautious.  

```{r}
# Sentence and police

#foo3 = lm(crmrte ~ polpc + avgsen, data = crime)
#foo3$coefficients
#vcov(foo3)

foo3log = lm(logcrmrte ~ polpc + avgsen, data = crime)
foo3log$coefficients
#vcov(foo3log)

foo3rows = c("logcrmrte", "crmrte", "polpc", "avgsen")

round(cor(crime[foo3rows]), 3)
scatterplotMatrix(crime[,foo3rows], diagonal = "histogram")


```

### 5. Economics
+ taxpc
+ wcon
+ wtuc
+ wtrd
+ wfir 
+ wser 
+ wmfg 
+ wfed 
+ wsta 
+ wloc

There's a lot to take in, however the negative relatinship to wser (wage service worker) is initially the most interesting. 

```{r}
# Economics
#foo4 = lm(crmrte ~ taxpc + wcon + wtuc + wtrd + wfir + wser + wmfg + wfed + wsta + wloc, data = crime)
#foo4$coefficients
#vcov(foo4)

foo4log = lm(logcrmrte ~ taxpc + wcon + wtuc + wtrd + wfir + wser + wmfg + wfed + wsta + wloc, data = crime)
foo4log$coefficients
#vcov(foo4log)

foo4rows = c("logcrmrte", "crmrte", "taxpc", "wcon", "wtuc", "wtrd", "wfir", "wser", "wmfg", "wfed", "wsta", "wloc")

round(cor(crime[foo4rows]), 2)
#scatterplotMatrix(crime[,foo4rows], diagonal = "histogram")

```



### 6. Demographics
+ pctmin80 
+ pctymle

pctmin80 is positively correlated and mix is negatively correlated.  This is counter intuitive to me.  Requires further

```{r}
# Demographics
#foo5 = lm(crmrte ~ pctmin80 + mix + pctymle, data = crime)
#foo5$coefficients
#vcov(foo5)

foo5log = lm(logcrmrte ~pctmin80 + mix + pctymle, data = crime)
foo5log$coefficients
#vcov(foo5log)

foo5rows = c("logcrmrte", "crmrte", "pctmin80", "pctymle")

round(cor(crime[foo5rows]), 3)
scatterplotMatrix(crime[,foo5rows], diagonal = "histogram")

```

### 7. Crime types

The higher the ratio of face-to-face crimes ends up with fewer crimes.  I suspect this is the result of a small police force that doesn't have as much time to go after less significant crimes, so I added that variable in too.  They're not strongly correlated.

```{r}
# Crime Types
foo6log = lm(logcrmrte ~ mix + polpc, data = crime)
foo6log$coefficients
#vcov(foo5log)

foo6rows = c("logcrmrte", "crmrte", "mix", "polpc")

round(cor(crime[foo6rows]), 3)
scatterplotMatrix(crime[,foo6rows], diagonal = "histogram")

```



```{r} 
plot(crime$crmrte, ylab = 'crimes committed per person')
```

```{r}
hist(crime$crmrte, xlab = 'crimes committed per person', main = 'Histogram of crimes committed per person')
```

```{r}
model1 <- lm(crmrte ~ prbarr + polpc + density, data = crime)
(model1$coefficients)

```


## Steps for evaluating variables
Leverage (and Influence if required)  
Goodness-of-Fit : AIC  
Omitted variable bias  
MSE  
E[theta hat] = theta

```{r}
crime$urban + crime$west + crime$central
crime$urban + crime$west
crime$urban + crime$central
crime$west + crime$central

```

